% from http://code.google.com/p/latex-byu-thesis


\usepackage[
    pdftex,
    bookmarks=true,
    bookmarksnumbered=true,
    breaklinks=false,
    raiselinks=true,
    pdfborder={0 0 0},
    colorlinks=false,
    bookmarksnumbered=true,
    ]{hyperref}

\usepackage[all]{hypcap}
\usepackage{natbib} % might not need this

% subfigures like 4.1.2
%\renewcommand{\thesubfigure}{\thefigure.\arabic{subfigure}}
%\makeatletter
%\renewcommand{\p@subfigure}{}
%\renewcommand{\@thesubfigure}{\thesubfigure:\hskip\subfiglabelskip}
%\makeatother

\setcounter{lofdepth}{2}

\newcommand{\Title}{Automatic Transition To Peer-to-Peer Download}
\newcommand{\Author}{Roger Deloy Pack}
\newcommand{\MonthGraduated}{April} % August
\newcommand{\YearGraduated}{2010}

\hypersetup{%
    pdftitle=\Title,%
    pdfauthor=\Author,%
    pdfsubject={MS Thesis, BYU CS Department: %
                Degree Granted \MonthGraduated~\YearGraduated, Document Created \today},%
    pdfkeywords={BYU, thesis, dissertation, LaTeX},%
}

% Rewrite the itemize, description, and enumerate environments to have more
% reasonable spacing:
\newcommand{\ItemSep}{\itemsep 0pt}
\let\oldenum=\enumerate
\renewcommand{\enumerate}{\oldenum \ItemSep}
\let\olditem=\itemize
\renewcommand{\itemize}{\olditem \ItemSep}
\let\olddesc=\description
\renewcommand{\description}{\olddesc \ItemSep}

% Important settings for the byumsphd class.
\title{\Title}
\author{\Author}

\committeechair{Daniel~Zappala}
\committeemembera{Mark~J.~Clement}
\committeememberb{Christophe~Giraud-Carrier}

\monthgraduated{\MonthGraduated}
\yeargraduated{\YearGraduated}
\yearcopyrighted{\YearGraduated}

\documentabstract{
For traditional web servers, available bandwidth decreases as the number of clients increases.  
This can cause servers to serve files slowly or to become completely overwhelmed when load grows too high.
BitTorrent is a peer-to-peer solution to this problem, but it requires manual configuration for each
file to be delivered this way.
We develop a new system that integrates peer-to-peer file delivery with traditional client-server downloads.  
Clients initially attempt to download a file from a web server; if this is too slow, they transition to peer-to-peer
delivery.  Experiments with a prototype system show that it serves up to 30x faster than traditional web servers.
}

\documentkeywords{%
    Peer-to-Peer Web, BitTorrent
}

\acknowledgments{%
    Thank you to my loving wife, my Mom, and my other family members who have made this possible.  You deserve to read this!
}

\department{Computer~Science}
\graduatecoordinator{Kent~E.~Seamons}
\collegedean{Thomas~W.~Sederberg}
\collegedeantitle{Associate~Dean}

% Remove all widows an orphans.  This is not normally recommended, but in a
% paper dissertation there is no reasonable way around it; you can't exactly
% rewrite already-published content to fix the problem.
\clubpenalty 10000
\widowpenalty 10000

% Allow pages to have extra blank space at the bottom in order to accommodate
% removal of widows and orphans.
\raggedbottom

% Get a little less fussy about word spacing on a line.  Sometimes produces
% ugly results, so keep your eyes peeled.
\sloppy